The experiments and evaluations conducted in this study demonstrate the trade-offs inherent in optimizing
lightweight models for practical tasks like typing assistance and summarization.
While our current optimizations and pipelines have shown measurable improvements in model size and inference time,
several avenues remain to enhance performance and usability further.

One promising direction is exploring frameworks such as ONNX, which can streamline operations like INT8 matrix multiplication
and improve runtime compatibility across diverse hardware.
Integrating ONNX into our workflow could significantly reduce inference latency for quantized models.
Another area of improvement lies in better utilization of PyTorch's capabilities.
While Python's inherent performance overhead has been mitigated to some extent by PyTorch's C++ backend,
leveraging Just-In-Time (JIT) compilation and converting models to TorchScript can further eliminate Python-specific bottlenecks.
This approach would ensure faster execution during inference without requiring extensive changes to existing training workflows.
For fine-tuning, emerging techniques like Low-Rank Adaptation (LoRA) offer a potential path forward.
By splitting weight matrices into low-rank components,
LoRA reduces the computational cost of gradient updates during fine-tuning while maintaining high precision.
Applying such methods to the LSTM and GPT-2 models may lead to faster convergence and better resource utilization.
Finally, future efforts should investigate optimizing the hybrid system itself.
For example, dynamic task allocation mechanisms could be developed to improve the decision-making process between LSTM and GPT-2,
ensuring optimal resource use based on input complexity.
Additionally, refining the LSTM to handle broader contexts or partially fine-tuning GPT-2 on character-level tasks
might further enhance the system's adaptability.

In conclusion, advancing lightweight Transformer models requires a balanced approach,
combining innovative algorithmic strategies with practical implementation refinements.
By leveraging emerging technologies and techniques, the potential for creating even more efficient and capable systems is substantial.
